---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->  

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.path = "README-", 
                      eval = FALSE)
```

# pbr

## Description

This is the companion R Package to the book *Plant Breeding in R*. It is primarily used to load all of the packages necessary for completing the tasks outlined in *Plant Breeding in R*. However, there are some functions that are provided within the package that may be useful.

## Installation

You can install this package using the `devtools` package:

```{r gh-install}
devtools::install_github("neyhartj/pbr")
```

When you install `pbr`, a number of other packages are installed. The **core** set of packages - those that you might need for most analyses - include:

- `dplyr`, `tidyr`, `readr`, for data management and modeling
- `ggplot2`, for visualization
- `lme4`, for fitting mixed-effect models
- `agridat`, for various agricultural experiment data

Additionally, other packages are installed that may be useful for specialized analysis, including those for:

- Managing genetic experiments and QTL mapping

    + `qtl`, for QTL mapping and simulations
    + `rrBLUP`, for GWAS and genomic selection
    
- Statistics:

    + `agricolae`, for statistical procedures in agriculture research
    + `BGLR`, for Bayesian linear regression
    + `lsmeans`, for calculating least-square means
    
    
## Usage
  
Loading the `pbr` package also loads the **core** set of packages (dplyr, tidyr, readr, ggplot, lme4, agridat):

```{r lib-pbr, eval=TRUE}
library(pbr)
```

Packages that are not in the core set must be loaded individually using the `library()` function, for instance:

```{r lib-qtl}
library(qtl)
```


## Functions

The `pbr` package includes endogenous functions to support the other packages. Those include

  Function           Purpose
------------------- ---------------------------------------------------------------
  `dist_env()`       Clustering of environments based on phenoypic observations
  `herit()`          Estimating heritability and variance components from fitted model object
  `herit_boot()`     Calculating confidence intervals for heritability based on bootstrapping

## Support

Please [open an issue](https://github.com/neyhartj/pbr/issues/new) for support or to comment.
